Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2017-09-05 redaction/Presentations/DEGIT_Conference/slides_DEGIT.tex
Nouvelle version des slides DEGIT

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] Part of the ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\begin{itemize}
\item[-] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
\footnotesize
$$pq = \tau \widetilde{p}q$$
\normalsize
\end{itemize}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
\begin{itemize}
\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
\begin{itemize}
\item[$\star$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2014) \vspace{0.1cm}
\end{itemize}
\item[-] Some trade policy instruments as well (quota license price)
\footnotesize
$$pq = (t+\widetilde{p})q  $$
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Our paper in one question (and 3 answers)}
\begin{itemize}
\item \textbf{Do additive transport costs matter?} \\ \vspace{0.2cm}
\item[$\Rightarrow$] \textbf{Yes, they do} \vspace{0.1cm}
\begin{itemize}
\item[-] An empirical decomposition of the structure of transport costs \vspace{0.1cm}
\item[-] Using the US imports database over 1974-2013 (air/vessel)\vspace{0.1cm}
\end{itemize}
\item[(1)] \textbf{In terms of size}: Additive costs are sizeable \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel \vspace{0.1cm}
%}
\end{itemize}
\item[(2)] \textbf{In terms of quality of fit}: With the additive component included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' %}
%}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{More on our database}
\begin{itemize}
\item Implications (and limitations) \vspace{0.1cm}
\begin{itemize}
\item[-] Only cover international transport costs \vspace{0.1cm}
\item[-] Among transport costs, insurance + handling + quantitative freight costs (e.g. not those related to the time value of goods) \vspace{0.1cm}
%\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
\end{itemize}
\item A rich database to exploit \vspace{0.1cm}
\begin{itemize}
\item[-] US imports, long time period: Broad view of international trade flows \vspace{0.1cm}
\item[-] A reliable database, already used by Hummels (2007), but on a shorter length of time \vspace{0.1cm}
\item[-] Have both the import and the export prices  \vspace{0.1cm}
\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive trade costs  \vspace{0.1cm}
\begin{itemize}
\item[$\neq$] Irrarazabal et al. (2015)
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The equation at the root of the estimation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_estimation_method]
\begin{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1>0$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation*}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
\end{equation*}
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
\end{itemize}
\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_estimation_method}{\beamergotobutton{More}}\vspace{0.1cm}

@@

\begin{frame}[label = slide_compeffects_strategy]
\frametitle{Eliminating composition effects: Our strategy}
\begin{itemize}
\item Based on our estimates of additive/multiplicative transport costs (TC) \hyperlink{app_compeffects_strategy}{\beamergotobutton{Comparison with Hummels (2007)}}\vspace{0.1cm}
\begin{itemize}
\item[-] That vary over time ($t$)/product ($k$) /origin country ($i$) (by transport mode): $\widehat{\tau}_{ikt}, \widehat{t}_{ikt}$ \vspace{0.1cm}
\end{itemize}
\item Starting from these values, extract the ``pure'' TC component \hyperlink{app_compeffects_details1}{\beamergotobutton{More details}}\vspace{0.1cm}
\begin{itemize}
\item[-] By the mean of a time fixed-effect \vspace{0.1cm}
\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted \vspace{0.1cm}
%\item[-] From which we extract the fitted ``total'' transport cost measure \vspace{0.1cm}
\end{itemize}
\item What explains the decrease of transport costs over time?
\begin{itemize}
\item[-] Is it mostly due to the composition of trade (product quality/weight, less distant countries of origin...) \vspace{0.1cm}
\item[-] or to the ``pure'' transports costs dimension (e.g., handling costs) ?
\end{itemize}
\item[$\Rightarrow$] Compare the unfitted TC component to the fitted one (composition effects excluded) \vspace{0.1cm}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Time trends in transport costs: Comments}
\begin{itemize}
\item International transport costs have substantially reduced over the period (1974-2013) \vspace{0.1cm}
\begin{itemize}
\item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
\item[-] Consistent with the literature, see Lafourcade \& Thisse (2011) \vspace{0.1cm}
\item[-] Roughly same order of magnitude for both the ad-valorem and the additive component  \vspace{0.1cm}
\end{itemize}
\item \textbf{Not much evidence of a difference between the ``raw'' and the ``pure'' TC components }\vspace{0.1cm}
\item In Air transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 50\% decrease, mainly attributable to the reduction of ``pure'' transport costs \vspace{0.1cm}
\item[-] For both the additive and the ad-valorem components  \vspace{0.1cm}
\end{itemize}
\normalsize
\item Composition effects (slightly) more pronounced in maritime transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 60\% decrease: Roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
\item[-] But for the multiplicative component only \vspace{0.1cm}
\end{itemize}
\normalsize
\end{itemize}
\end{frame}

>>>>

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] Part of the ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
\item[$\Rightarrow$] Multiplicative costs:     \footnotesize $pq = \tau \widetilde{p}q$
\normalsize
\end{itemize}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
\begin{itemize}
\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2015) \vspace{0.1cm}
\item[$\ast$] Some trade policy instruments as well (quota license price) \vspace{0.1cm}
\item[$\Rightarrow$] Additive costs: \footnotesize $pq = (t+\widetilde{p})q  $
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Our paper in one question (and 3 answers)}

\begin{itemize}
\item  \textbf{Do additive transport costs matter?} \vspace{0.1cm}
\begin{itemize}
\item[-] Provide an empirical decomposition of the structure of transport costs \vspace{0.1cm}
\item[-] Using the US imports database over 1974-2013 (air/vessel)\vspace{0.2cm}
\end{itemize}
\item[$\Rightarrow$] \textbf{Yes, they do} \vspace{0.1cm}

\item[(1)] \textbf{In terms of size}: Additive costs are sizeable \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel \vspace{0.1cm}
%\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Roughly 50\% of the overall transport costs \vspace{0.1cm}

%}
\end{itemize}
\item[(2)] \textbf{In terms of quality of fit}: With the additive component included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' %}
%}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{More on our database}
\begin{itemize}
\item Implications (and limitations) \vspace{0.1cm}
\begin{itemize}
\item[-] Only cover international transport costs \vspace{0.1cm}
\item[-] Among transport costs, insurance + handling + quantitative freight costs (e.g. not those related to the time value of goods) \vspace{0.1cm}
%\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
\end{itemize}
\item A rich database to exploit \vspace{0.1cm}
\begin{itemize}
\item[-] US imports, long time period: Broad view of international trade flows \vspace{0.1cm}
\item[-] A reliable database, already used by Hummels (2007), but on a shorter length of time \vspace{0.1cm}
\item[-] Have both the import and the export prices  \vspace{0.1cm}
\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive transport costs  \vspace{0.1cm}
\begin{itemize}
\item[$\neq$] Irrarazabal et al. (2015)
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The equation at the root of the estimation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:basis_equation}
\end{equation}
\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_estimation_method]
\begin{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} >1$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest (\ref{eq:basis_equation}) becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation*}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
\end{equation*}
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
\end{itemize}
\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_estimation_method}{\beamergotobutton{More}}\vspace{0.1cm}

@@

\begin{frame} [label = slide_compeffects_strategy]
\frametitle{Eliminating composition effects: Our strategy}
\begin{itemize}
\item Based on our estimates of additive/multiplicative transport costs (TC) \hyperlink{app_compeffects_strategy}{\beamergotobutton{Comparison with Hummels (2007)}}\vspace{0.1cm}
\begin{itemize}
\item[-] That vary over time ($t$)/product ($k$) /origin country ($i$) (by transport mode): $\widehat{\tau}_{ikt}, \widehat{t}_{ikt}$ \vspace{0.1cm}
\end{itemize}
\item Starting from these values, extract the ``pure'' TC component \hyperlink{app_compeffects_details1}{\beamergotobutton{More details}}\vspace{0.1cm}
\begin{itemize}
\item[-] By the mean of a time fixed-effect \vspace{0.1cm}
\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted \vspace{0.2cm}
%\item[-] From which we extract the fitted ``total'' transport cost measure \vspace{0.1cm}
\end{itemize}
\item What explains the decrease of transport costs over time?  \vspace{0.1cm}
\begin{itemize}
\item[-] Composition effects vs ``Pure'' TC changes? \vspace{0.1cm}
%\item[-] Is it mostly due to the composition of trade (product quality/weight, less distant countries of origin...) \vspace{0.1cm}
%\item[-] or to the ``pure'' transports costs dimension (e.g., handling costs) ?
\end{itemize}
\item[$\Rightarrow$] Compare the unfitted TC component to the fitted one (composition effects excluded) \vspace{0.1cm}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Time trends in transport costs: Comments}
\begin{itemize}
\item A strong reduction of transport costs over the period (1974-2013) \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
\item[-] Consistent with the literature, see Lafourcade \& Thisse (2011) \vspace{0.1cm}
\item[-] Roughly same order of magnitude for both the ad-valorem and the additive component  \vspace{0.1cm}
\normalsize
\end{itemize}
\item \textbf{Trade composition effects: A minor role in accounting for the time trend of TC}\vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] Not much evidence of a difference between the ``raw'' and the ``pure'' TC components \vspace{0.1cm}
\normalsize
\end{itemize}
\item In Air transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 50\% decrease, mainly attributable to the reduction of ``pure'' transport costs \vspace{0.1cm}
\item[-] For both the additive and the ad-valorem components  \vspace{0.1cm}
\end{itemize}
\normalsize
\item Composition effects (slightly) more pronounced in maritime transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 60\% decrease: Roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
\item[-] But for the multiplicative component only \vspace{0.1cm}
\end{itemize}
\normalsize
\end{itemize}
\end{frame}

### 2017-09-05 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex
Nouvelle version des slides DEGIT

\begin{frame}
\frametitle{Our paper in 3 questions (and 3 answers)}
An empirical decomposition of the structure of transport costs \vspace{0.2cm}
\begin{itemize}
\item[(1)]\textbf{ What is the size of the iceberg and the additive costs?} \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.2cm}
%}
\end{itemize}
\item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
%}
\end{itemize}
\end{itemize}
\end{frame}

>>>>

\begin{frame}
\frametitle{Our paper in 3 questions (and 3 answers)}
An empirical decomposition of the structure of transport costs \vspace{0.2cm}
\begin{itemize}
\item[(1)]\textbf{ What is the size of the iceberg and the additive costs?} \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both, using US imports data \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] Additive cost: 1.8\% and 2.9\% of the export price, in air / vessel  \vspace{0.1cm}
\item[-] Iceberg cost: 2.5\% of the export price in air, and 3.2\% in vessel (mean value over 1974-2013) \vspace{0.2cm}

%}
\end{itemize}
\item[(2)] \textbf{ What do we lose by skipping the additive part of transport costs?} \vspace{0.1cm}
\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
%}
\end{itemize}
\end{itemize}
\end{frame}
